Warming non-randomly speeds up food web fluxes in both an absolute and relative sense

### 2022-07-20 doc/ms/ms.Rmd
updated figures and added some discussion text

Rising global temperatures are altering the pathways of energy and material flow across and within ecosystems with potential consequences for the services they provide. While a growing body of theoretical and empirical study has enhanced our knowledge of temperature-mediated changes to ecosystems [\ e.g., @oconnor2009], general patterns are uncertain and empirical studies often idiosyncratic [@zhang2017; @nelson2017]---especially at higher levels of organizations such as communities and food webs [@walther2002; @woodward2010]. Here, we document shifts in the pathways of material flow  across a wide natural temperature gradient and show that increasing temperatures were associated with reductions in mean population body size and increases in mean population *P:B* across stream consumer communities. Further, increasing temperatures also systematically skewed OM fluxes *within* communities such that, food web fluxes were increasingly skewed towards smaller, quick turnover populations at warmer temperatures. Lastly, there was evidence that the distribution of fluxes within communities was non-randomly organized, especially at temperature extremes and in relation to population *P:B*, suggesting that environmental filtering is especially important in community assembly of both cold and warm communities. This tendency for energy fluxes to be dominated by relatively higher turnover populations at warmer temperatures and this organization to be increasingly non-random, suggests the speeding up of energy and material fluxes through ecosystems in both an absolute and relative sense may be a general effect of increasing temperature.

@@

Temperature has long been known to be an important control on ecological process rates [@arrhenius1897; @gillooly2003; @brown2004]

>>>>

Rising global temperatures are altering the pathways of energy and material flow across and within ecosystems with potential consequences for the services they provide. While a growing body of theoretical and empirical study has enhanced our knowledge of temperature-mediated changes to ecosystems [\ e.g., @oconnor2009], general patterns are uncertain and empirical studies often idiosyncratic [@zhang2017; @nelson2017]---especially at higher levels of organizations such as communities and food webs [@walther2002; @woodward2010]. Here, we document shifts in the pathways of material flow  across a wide natural temperature gradient and show that increasing temperatures were associated with reductions in mean population body size and increases in mean population *P:B* across stream consumer communities. Further, increasing temperature also systematically skewed OM fluxes *within* communities such that, food web fluxes were increasingly skewed towards smaller, quick turnover populations at warmer temperatures. Lastly, there was evidence that the distribution of fluxes within communities was non-randomly organized, especially at moderate to warm temperatures and in relation to population *P:B*, suggesting that environmental filtering is especially important in community assembly of warm communities. This tendency for energy fluxes to be dominated by relatively higher turnover populations at warmer temperatures and this organization to be increasingly non-random, suggests the speeding up of energy and material fluxes through ecosystems in both an absolute and relative sense may be a general effect of increasing temperature.

Reductions in organism body size with warming has been deemed a "universal" response to warming [@daufresnes2009; @gardner2011; @uszko2022]. We observed a clear decrease in the average individual body size of populations from cool to warm communities (Figure 2), corresponding to ~9% decline in the mass of individuals for every 1$^\circ$C increase in temperature. While this change represents a community-level shift, the magnitude of change is consistent with recent work on the effects of warming on organism body size across broad taxonomic groups [@deutsch2022]. This reduction in body size with increasing temperature contrasts results from previous whole-ecosystem warming experiments which observed shifts in community biomass and productivity from smaller to larger organisms following a 3$^\circ$C increase in temperature [@nelson2017; @nelson2017a]. Our results also counter observational studies in this system that have shown no consistent change in body size within diatom communities at either the population- or community-scale [@adams2013]. These contrasting results are consistent with deviations from the "universal" response observed across taxonomic groups [e.g., diatoms, @adams2013; invertebrates @zeuss2017; birds @geist1987; fish @rypel2014]. Explanations for these deviations vary [@ohlberger2013], some highlight additional processes such as competition and predation [@delong2014], while others propose alternative models of organism body size optimization, such as the balance of resource supply and energy demand [e.g., @kozlowski2004; @delong2012a]. Crucially, these alternative models highlight the trade-off between asymptotic body size and metabolic demand [@delong2012a] that can shape life history patterns (i.e., voltinism, @zeuss2017; @nelson2020b). Considering the connections between organism body size and many other ecological attributes [@peters1983], reductions in body size with increasing temperatures could have broad implications for ecosystem dynamics with a changing climate [@gibert2019a].

Warmer ecosystems exhibited higher rates of energy flux in both an absolute and relative sense. At the community level, we observed a ~7% increase in the mean biomass turnover rate (P:B; y^-1^). Temperature has long been recognized as an important driver of ecological rates [@gillooly2001] and the increase in biomass turnover among communities is likely attributed to the thermodynamic effect of temperature--in part. Biomass turnover rate is also tightly tied to organism body size [@brown2004; @huryn2007] suggesting that temperature may influence the distribution of *P:B*s both directly through its effects on metabolic rates and indirectly by inducing reductions in organism body size. The influence of this interaction between the direct and indirect effect of temperature on metabolic activities in increasingly recognized. For example, the effects of temperature on population carrying capacity appear to deviate from metabolic scaling predictions until the effects of concomitant changes to organism body size are taken into account [@bernhardt2018]. Further, reduce body sizes stabilize systems by reducing C:R ratios at higher temperatures [@gilbert; @sentis2017; @bideault2019]

@@

Temperature has long been known to be an important control on ecological process rates [@gillooly2003; @brown2004]
